A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-11-25 kobakian-cook.Rmd
conclusions

This paper provides an algorithm for Australia called the hexagon tile map.
It also describes how this algorithm addresses the potential issues found with contiguous, non-contiguous and Dorling cartograms when applied to Australia. It achieves a display that effectively presents the density of the population in small geographic areas and de emphasises the large rural geographic space between the densely populated capital cities.

<!-- implementation -->
The hexagon tile map display acknowledges that the amount of residents can differ but each administrative area is equally important. This solution uses equally sized areas, and maintain neighbourhood boundary connections. The `sugarbag` package for `R` automates the creation of tessellated hexagon tile maps by providing an algorithm to design these displays. The Australian application preserves spatial relationships, and emphasises capital cities. 
However, similar to the choropleth map display the tessellation does not allow the size of the hexagons to represent another variable. The algorithm is heavily dependent on the focal points used, as this determines the order of allocation. With careful consideration of the choropleth map, the small geographic inner city areas may not have been noticed by viewers, but the hexagon tile map display emphasises them. The communities in northern Tasmania and the Northern territory do not draw attention because of their size as in the choropleth, but their colour is still noticeably below average when contrasted with the hexagons in New South Wales.

<!-- limitiations, where it might not work -->

This algorithm should be general enough to apply to other study regions and contexts. The algorithm has only been tested using single countries. While the buffer allows extension beyond the furthest centroids, there is no mechanism implemented to protect the borders and ensure centroids are placed within the geographic borders of a country.
The algorithm is also currently limited to representing each geographic area with only one hexagon. Future developments may address these limitations, and refine the filter steps for final selection of a hexagon grid point for each centroid. A direct angle is currently implemented, but may narrowly miss appropriate hexagons, a logarithmic function may help to choose a closer hexagon to the original centroid location.
This algorithm is an effective start to creating hexagon tile maps for many geographic units. This contributes to an extensive body of work that encourages the use of alternative map displays for statistical displays of spatial data.

>>>>

This work provides an algorithm called the hexagon tile map.
It also describes how this algorithm addresses the potential issues found with contiguous, non-contiguous and Dorling cartograms when applied to Australia. It achieves a display that effectively presents the density of the population in small geographic areas and de-emphasises the large rural geographic space between the densely populated capital cities. The hexagon tile map display acknowledges that the amount of residents can differ but each administrative area is equally important. This solution uses equally sized areas, and maintain neighbourhood boundary connections.

<!-- implementation -->
 The `sugarbag` package for `R` automates the creation of tessellated hexagon tile maps by providing an algorithm to design these displays. The Australian application preserves spatial relationships, and emphasises capital cities. However, similar to the choropleth map display the tessellation does not allow the size of the hexagons to represent another variable. The algorithm uses focal points, which are the centers of capital cities for Australia, to determine the order and direction of allocation. After careful consideration of the choropleth map, the small geographic inner city areas with high thyroid cancer prevalence may have been missed by viewers, but the hexagon tile map display reveals them. The communities in northern Tasmania and the Northern Territory draw less attention because they are smaller and more isolated in the hexagon tile map, but their colour is still comparable with the hexagons in other regions.

<!-- limitations, where it might not work -->

This algorithm should be general enough to apply to other study regions and contexts. The algorithm has only been tested using single countries. While the buffer allows extension beyond the furthest centroids, there is no mechanism implemented to protect the borders and ensure centroids are placed within the geographic borders of a country.
The algorithm is also currently limited to representing each geographic area with only one hexagon. Future developments may address these limitations, and refine the filter steps for final selection of a hexagon grid point for each centroid. A direct angle is currently implemented, but may narrowly miss appropriate hexagons, a logarithmic function may help to choose a closer hexagon to the original centroid location.
This algorithm is an effective start to automatically creating hexagon tile maps for many geographic units. This contributes to an extensive body of work that encourages the use of alternative map displays for statistical displays of spatial data.

### 2020-11-25 kobakian-cook.tex
conclusions

Administrative areas are often used to aggregate census data and used to
understand demographics within communities of the Australian population.
Each SA2 \citep{abs2011} was designed to represent a community. This
collection of communities presents an opportunity for explicit map
transformations to improve communication of spatial distributions
\citep{CBATCC}. In figure \ref{fig:tas_displays} the choropleth map can
be seen underneath each alternative map display to allow for comparisons
to be made.

@@

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The connected communities can be
recognised by allowing the boundaries to maintain connection in the
transformed display. The contiguous cartogram displayed in figure
\ref{fig:tas_displays}a) draws attention to smaller geographic units
when they are rescaled according to the population \citep{DMAHP}. These
new shapes can now be coloured to represent a second variable. The
algorithm uses the geographic shape of the areas and iterates toward
sizing the areas to represent the population. This display can create
twisted and unfamiliar shapes from the geographic units as the
algorithms must satisfy the topology conditions, especially when there
are communities located geographically far from their neighbours
\citep{TVSSS}.

The non-contiguous cartogram in figure \ref{fig:tas_displays}b) also
uses the population to rescale the geographic units. Unlike the
contiguous cartogram, the SA2 areas maintain their geographic shape, but
they may not retain the connection with their neighbours. The population
of the SA2 areas is used to scale the geographic units in the
non-contiguous cartogram \citep{NAC}. The amount of background space can
be meaningful in non-contiguous cartograms \citep{ECGC}. The size
disparity between rural areas and urban areas result in reasonable
display within the south eastern city of Hobart, but these units are not
visible in the context of the Tasmania state map. Depending on the
difference between the population and geographic land size, the amount
of white space can also prevent meaningful understanding of the
distribution \citep{TVSSS}.

@@

The \texttt{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in figure
\ref{fig:filterprocess}c.~This allows the spatial relationship to be
preserved, even when it is allocated to a hexagon that is further from
the focal point then the original centroid location.

@@

This paper provides an algorithm for Australia called the hexagon tile
map. It also describes how this algorithm addresses the potential issues
found with contiguous, non-contiguous and Dorling cartograms when
applied to Australia. It achieves a display that effectively presents
the density of the population in small geographic areas and de
emphasises the large rural geographic space between the densely
populated capital cities.

The hexagon tile map display acknowledges that the amount of residents
can differ but each administrative area is equally important. This
solution uses equally sized areas, and maintain neighbourhood boundary
connections. The \texttt{sugarbag} package for \texttt{R} automates the
creation of tessellated hexagon tile maps by providing an algorithm to
design these displays. The Australian application preserves spatial
relationships, and emphasises capital cities. However, similar to the
choropleth map display the tessellation does not allow the size of the
hexagons to represent another variable. The algorithm is heavily
dependent on the focal points used, as this determines the order of
allocation. With careful consideration of the choropleth map, the small
geographic inner city areas may not have been noticed by viewers, but
the hexagon tile map display emphasises them. The communities in
northern Tasmania and the Northern territory do not draw attention
because of their size as in the choropleth, but their colour is still
noticeably below average when contrasted with the hexagons in New South
Wales.

>>>>

Administrative areas are often used to aggregate census data and used to
understand demographics within communities of the Australian population.
Each SA2 \citep{abs2011} was designed to represent a community. This
collection of communities presents an opportunity for explicit map
transformations to improve communication of spatial distributions
\citep{CBATCC}. In Figure \ref{fig:tas_displays} the choropleth map can
be seen underneath each alternative map display to allow for comparisons
to be made.

@@

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The connected communities can be
recognised by allowing the boundaries to maintain connection in the
transformed display. The contiguous cartogram displayed in Figure
\ref{fig:tas_displays}a) draws attention to smaller geographic units
when they are rescaled according to the population \citep{DMAHP}. These
new shapes can now be coloured to represent a second variable. The
algorithm uses the geographic shape of the areas and iterates toward
sizing the areas to represent the population. This display can create
twisted and unfamiliar shapes from the geographic units as the
algorithms must satisfy the topology conditions, especially when there
are communities located geographically far from their neighbours
\citep{TVSSS}.

The non-contiguous cartogram in Figure \ref{fig:tas_displays}b) also
uses the population to rescale the geographic units. Unlike the
contiguous cartogram, the SA2 areas maintain their geographic shape, but
they may not retain the connection with their neighbours. The population
of the SA2 areas is used to scale the geographic units in the
non-contiguous cartogram \citep{NAC}. The amount of background space can
be meaningful in non-contiguous cartograms \citep{ECGC}. The size
disparity between rural areas and urban areas result in reasonable
display within the south eastern city of Hobart, but these units are not
visible in the context of the Tasmania state map. Depending on the
difference between the population and geographic land size, the amount
of white space can also prevent meaningful understanding of the
distribution \citep{TVSSS}.

@@

The \texttt{width} parameter is used to take a slice of the remaining
points. The width is the amount of degrees used on either side of the
angle from the focal point to centroid location. This uses the angle
from the closest capital city, to the current centroid as seen in Figure
\ref{fig:filterprocess}c.~This allows the spatial relationship to be
preserved, even when it is allocated to a hexagon that is further from
the focal point then the original centroid location.

@@

This work provides an algorithm called the hexagon tile map. It also
describes how this algorithm addresses the potential issues found with
contiguous, non-contiguous and Dorling cartograms when applied to
Australia. It achieves a display that effectively presents the density
of the population in small geographic areas and de-emphasises the large
rural geographic space between the densely populated capital cities. The
hexagon tile map display acknowledges that the amount of residents can
differ but each administrative area is equally important. This solution
uses equally sized areas, and maintain neighbourhood boundary
connections.

The \texttt{sugarbag} package for \texttt{R} automates the creation of
tessellated hexagon tile maps by providing an algorithm to design these
displays. The Australian application preserves spatial relationships,
and emphasises capital cities. However, similar to the choropleth map
display the tessellation does not allow the size of the hexagons to
represent another variable. The algorithm uses focal points, which are
the centers of capital cities for Australia, to determine the order and
direction of allocation. After careful consideration of the choropleth
map, the small geographic inner city areas with high thyroid cancer
prevalence may have been missed by viewers, but the hexagon tile map
display reveals them. The communities in northern Tasmania and the
Northern Territory draw less attention because they are smaller and more
isolated in the hexagon tile map, but their colour is still comparable
with the hexagons in other regions.
